Response to Reviewers

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Finalized new figure 4

# Per EEZs
top_toe <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
             ) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  summarise_at(vars(mean_yr),
               c("mean"=mean,
                 "sd" = sd)) %>% 
  mutate(
    geo_region = ifelse(is.na(geo_region),"Other",geo_region)
  )

# Numbers for paper
x <- top_toe %>% 
  filter(mean_yr <= 2021,
         variable == "value",
         tresh == selected_tresh)

length(unique(x$sau_fishing_entity))/length(unique(Sau_data$fishing_entity_id))

# Number of shifting stocks per geo-region
x %>% 
  group_by(geo_region) %>% 
  tally() %>% 
  arrange(n) %>% 
  top_frac(0.2,n) # top 20%

@@

mean(Top_map_data$mean)
sd(Top_map_data$mean)

World_map_top <- World_map %>%
  full_join(Top_map_data,
            by = c("sau_fishing_entity")
  ) %>% 
  filter(!is.na(variable))

# These are regions without transboundary species
grey_land <- World_map %>% 
  filter(!sau_fishing_entity %in% Top_map_data$sau_fishing_entity) %>% 
  filter(!sovereignt %in% c("Antarctica"))

@@

  map <- ggplot(World_map_top) +
    geom_sf(
      aes(
        fill = mean #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # grey maps
    # ggplot() +
    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
    # geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    # Set Participation labels
    scale_fill_viridis("Average year of range shift",
                       limits = c(2000,2100),
                       breaks = seq(2000,2100,25),
                       direction = -1
    ) +
    my_ggtheme_m(leg_pos = "bottom") +
    theme(legend.key.width = unit(5, "line")) +
    guides(
      fill=guide_colorbar(title.position="top", 
                        title.hjust =0.5)
  )

>>>>

top_stocks_per_country <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
  summarise(n()) %>% 
  left_join(Sau_data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(!is.na(variable)) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  # Select top 5 most valuable stocks
  top_n(5,value)

# Per EEZs
top_toe <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
             ) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  summarise_at(vars(mean_yr),
               c("mean"=mean,
                 "sd" = sd)) %>% 
  mutate(
    geo_region = ifelse(is.na(geo_region),"Other",geo_region)
  )

# Numbers for paper
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
             ) %>% 
  group_by(variable,tresh) %>% 
  summarise_at(vars(mean_yr),
               c("mean"=mean,
                 "sd" = sd))

# Number of coastal nations
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
             ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(variable,tresh) %>% 
  summarise(
    n_countries = length(unique(sau_fishing_entity))
  ) %>% 
  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id)))

# Number of shifting stocks per geo-region
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
             ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(geo_region) %>% 
  tally() %>% 
  arrange(n) %>% 
  top_frac(0.25,n) # top 25%

@@

World_map_top <- World_eez_sf %>% 
  left_join(Matching_names) %>% 
  full_join(Top_map_data,
            by = c("sau_fishing_entity")
  ) %>% 
  filter(!is.na(variable))

@@

  no_toe_eez <- World_eez_sf %>% 
    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
           !eez_name %in% World_map_top$eez_name) %>%  # Filter EEZs with ToE stocks
    rbind(russia_churchil)

  no_toe_land <- World_map %>% 
    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
           !name %in% World_map_top$name) # Filter EEZs with ToE stocks

  map <- ggplot(World_map_top) +
    geom_sf(
      aes(
        fill = mean #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # grey maps
    # ggplot() +
    geom_sf(data = World_map, aes(), fill = "grey50", color =  "white",size = 0.1) +
    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    # geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    # Set Participation labels
    scale_fill_viridis("Average year of range shift of top valued transboundary stocks",
                       limits = c(2000,2100),
                       breaks = seq(2000,2100,25),
                       direction = -1
    ) +
    my_ggtheme_m(leg_pos = "bottom") +
    theme(legend.key.width = unit(5, "line")) +
    guides(
      fill=guide_colorbar(title.position="top", 
                        title.hjust =0.5)
  )

### 2021-11-12 Scripts/Manuscript.Rmd
Finalized new figure 4

We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.). Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d. threshold (Fig. \@ref(fig:toeFigure3); above Fisheries >75th percentile). In this case, 321 stocks shifting in just eleven countries account for 4.2 billion USD (Fig. \@ref(fig:toeFigure3)). The number of transboundary stocks shifting from historical variability was highest in the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66) (Fig. \@ref(fig:toeFigure3)). The values are substantially higher when considering a more relaxed threshold. In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs.

>>>>

**While the total number of shifting stocks is important for fisheries management, having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting. Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (New Fig 4). On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two( 2028 $\pm$ 22 years under threshold one). Moreover, our projections suggest that these high value stocks are already shifting in 52 (27%) coastal nations under threshold two (n = 85 44% under threshold one), with the regions of Latin America and the Caribbean (n = 63), Sub-Saharan Africa (n = 59), South-eastern Asia  (n = 28), and Polynesia (n = 19) presenting the larger number of shifting stocks (25% of total high value stocks).**

<!-- We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.). Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d. threshold (Fig. \@ref(fig:toeFigure3); above Fisheries >75th percentile). In this case, 321 stocks shifting in just eleven countries account for 4.2 billion USD (Fig. \@ref(fig:toeFigure3)). The number of transboundary stocks shifting from historical variability was highest in the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66) (Fig. \@ref(fig:toeFigure3)). The values are substantially higher when considering a more relaxed threshold. In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs. -->

### 2021-11-12 Scripts/responses_rev_gcb.Rmd
Finalized new figure 4

- **Response:** The definition of "stock" will likely influence the results as it is a fundamental part of the analysis. Replicating this analysis at the species level will undoubtedly result in multiple countries sharing each species. For example, in the case of Atlantic Cod, this would be shared by US, Canada and Greenland (and maybe enven northern Europe). Choosing this approach invariably comes with the assumption that fishing Cod in, e.g., the Gulf of Maine, will have direct impact on e.g., Greenland fisheries. While this might be true for some few highly migratory species (e.g., tunas) we believe is a stretch for all other species in this analysis.

@@

- **Response:** Please see responses to the reviewer's main comments. We have defined a transboundary stock as follows;

@@

- **Response:** Thank you for pointing this out. In the original manuscript we were showing the percentage difference between the two numbers (i.e., $x = \frac{|45-63|}{|\frac{45+63}{2}|}*100$), hence the 33.3% difference. However, now we realize this can be misleading and have adopted a more straightforward computation. We also corrected the 63% typo (not 67%). It now reads:

@@

- **Response:** Thank you for your commnets on our fiugre 3. We have devided the responses by each point you make. Please see below

@@

  - **Response:**  **NOTE: I do not (fully) agree with this comment. yes, the figure is complex, but I do not think having different color scales will make it any less complex. I actually think it will make it even harder to read. Do you think we can counter-argue this comment?**

@@

  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region. **NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency (and make the reviewer happy)?**

@@

- **Response:** We intended to show this information in Figure 4. However, the message in the figure is not clear, as the reviewer pointed out in the next comment. We have modified figure 4 to show this as a figure.

@@

>>>>

- **Response:** The definition of "stock" will likely influence the results as it is a fundamental part of the analysis. Replicating this analysis at the species level will undoubtedly result in multiple countries sharing each species. For example, in the case of Atlantic Cod, this would be shared by US, Canada and Greenland (and maybe even northern Europe). Choosing this approach invariably comes with the assumption that fishing Cod in, e.g., the Gulf of Maine, will have direct impact on e.g., Greenland fisheries. While this might be true for some few highly migratory species (e.g., tunas) we believe is a stretch for all other species in this analysis.

@@

- **Response:** The definition is now clearly stated in the revised manuscript as follows;

@@

- **Response:** Thank you for pointing this out. In the original manuscript we were showing the percentage difference between the two numbers (i.e., $x = \frac{|45-63|}{|\frac{45+63}{2}|}*100$), hence the 33.3% difference. However, now we realize this can be misleading and have adopted a more straightforward computation that aligns with the reviewer's math. We also corrected the 63% typo (not 67%). It now reads:

@@

- **Response:** Thank you for your comments on our figure 3 and see below the responses to each point you make.

@@

  - **Response:**  **NOTE: I do not (fully) agree with this comment, Do you think we can counter-argue it? (We had the same type of plot in the Sci. Rep. paper). Yes, the figure is complex, but I do not think having different color scales will make it any less complex. I actually think it will make it even harder to read.**

@@

  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region . **NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency or the inverse? (and make the reviewer happy)?**

@@

- **Response:** We agree with the reviewer that showing the biggest “winners and losers” in terms of high-value species would be interesting. Thus, we have included a new figure and text in this section showing the year of range shift of the 5 most valuable stocks per country (see response to main comments).

@@

- **Response:NOTE THINKING ON CHANGEING FIGURES** Our original intention was to show the importance in terms of value, of shifting stocks. However, we agree that this figure is confusing and it's message was ambiguous. Moreover, we think that the reviewers suggestion of the most valuable species brings home a mor important message and have decided to remove this figure from the revised manuscript and include the "top 5 most valuable stocks" figure instead.
